---
  title: "test"
output: 
  flexdashboard::flex_dashboard:
  theme: bootstrap
runtime: shiny
---
  
```{r setup,include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tibbletime)
library(dygraphs)
library(magrittr)
library(xts)
```

```{r global,include=FALSE}
# generate data
set.seed(1)
dat <- data.frame(date = seq(as.Date("2018-01-01"), 
                             as.Date("2018-06-30"), 
                             "days"),
                  sex = sample(c("male", "female"), 181, replace=TRUE),
                  lang = sample(c("english", "spanish"), 181, replace=TRUE),
                  age = sample(20:35, 181, replace=TRUE))
dat <- sample_n(dat, 80)

prep_dat <- function(dat, grpvars) {
  group_syms <- rlang::syms(grpvars)
  
  dat_week <- dat %>% 
    mutate(week=lubridate::floor_date(date, unit="week")) %>%
    group_by(!!!group_syms) %>% 
    tally()
  
  # if there is a second grouping column, spread the counts
  if (grpvars[2] != "week") {
    dat_week <- dat_week %>% 
      spread(!!rlang::sym(grpvars[2]), n)
  }
  
  ret_dat <- xts(dat_week, order.by=dat_week$week)
  #week column is no longer needed
  ret_dat$week <- NULL
  ret_dat
}

choices <- list("All" = 1, "By Sex" = 2, "By Language" = 3)
```

Sidebar {.sidebar}
=====================================
  
  ```{r}
radioButtons("diss", label = "Disaggregation",
             choices = choices, 
             selected = 1)
```

Page 1
=====================================
  
  ```{r}
get_grpvars <- reactive(
  case_when(
    input$diss == "1" ~ c("week"),
    input$diss == "2" ~ c("week", "sex"),
    input$diss == "3" ~ c("week", "lang"))
)

main_text <- reactive(
  names(choices)[as.integer(input$diss)]
)

get_dat4plot <- reactive({
  grpvars <- get_grpvars()
  print(grpvars)
  prep_dat(dat, grpvars)
})

# plot
renderDygraph({
  dat4plot <- get_dat4plot()
  dygraph(dat4plot, main=main_text()) %>%
    dyRangeSelector() %>%
    dyOptions(useDataTimezone = FALSE,
              stepPlot = TRUE,
              drawGrid = FALSE,
              fillGraph = TRUE) 
}
)
```

My answer probably overlaps with @Jon Spring and @Simon.S.A.'s, but I think mine goes further than their in implementation details. 

I found your group counts calculation can be done much more economically - you basically just need to do a group tally, either by week alone or week + sex/lang. You can program the grouping variable(s) using `tidyeval` + `dplyr` in your function (as @Simon.S.A mentioned in his answer).

After having the appropriate group counts, `spread` your data if you have week + sex/lang grouping so that your tibble only has a week column + count(s) for plotting and you don't need to specify the `dySeries` for your dygraph. Finally, convert the tibble to `xts` right before passing to plot by dygraph.

Here is my complete code based on your sample code.